Sperm should evolve to make female meiosis fair.

### 2014-07-21 writeup_and_notes/GeneticsResub/male_dependent_drive.tex
yb modified discussion and fixed some refs

\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although this effect may be
        weaks as such
        lineages may only persist for a short time). Another
        hypothesis that our work suggests is we might
        expect that phylogenetic variation in the stage of female meiosis when
        fertilization occurs to play a role in the prevalance of
        female meiotic drive.
For example, in taxa where meiosis I is completed
by the egg pre-fertilization, including humans, centromere drive may
be more common as sperm based modifiers can not intercede than in
species where fertilization occurs during MI. 
Such hypotheses could be tested if the speed of centromere
turnover serves as a proxy for the efficacy of MI drive. }

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
\yb{Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[(for a review and evaluation of such
  hypotheses see][]{Mira1998}.
However, given that arrest does occur its timing relative to
fertilisation, various features of the interaction between sperm and
egg and divergence of these features across species  \citep{Karrbook:09} could be subject
to selection to conflict between female meiotic drivers and alleles
elsewhere in the genomes of sperm and eggs.
the mechanism by which it occur
mechanistic features, and evolutionary changes among taxa

various features of the
mechanism of fertilization 
however, our work does imply that the fertilization requirement does
not 
leave female meiosis open to sabotage by sperm collaborators, but
rather that  an 
evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }

>>>>

\gc{Our theory also encourages phylogenetic hypotheses concerning the relationship between the opportunity 
	for female meiotic drive and the requirement of fertilization for the completion of female meiosis. 
For example, we predict that a lower opportunity for female meiotic drive,
        e.g. an animal lineage with a history of high inbreeding
        or selfing, 
        may be accompanied by a relaxation of the requirement of fertilization for the
        completion of female meiosis (although 
   \yb{     opportunities to test this hypothesis may be limited because }
        %this effect may be weaks as such
        lineages may only persist for a short time). 
Our results also suggests that 
%Another   hypothesis that our work suggests is we might expect 
        that phylogenetic variation in the stage of female meiosis when
        fertilization occurs to influence the prevalence of
        female meiotic drive.
For example, in taxa where meiosis I is completed
by the egg pre-fertilization, including humans, centromere drive may
be more common as sperm based modifiers can not intercede than in
species where fertilization occurs during MI. 
Such hypotheses could be tested if the speed of centromere
turnover serves as a proxy for the efficacy of MI drive. }
\newline

@@

Our results highlight potentially counterintuitive results of complex genetic conflicts. 
Despite much opportunity for conflict between sperm and females over fertilization \citep{Partridge1998},
	the interests of fertilizing sperm and female are quite well aligned during syngamy.
While conflict between mother and her alternative chromosomes ensues, 
	fertilizing sperm decidedly side with mom, as both have a shared interest in producing a viable and potentially fit offspring. 
\yb{Our model does not directly speak to the evolutionary origin of
  female meiotic arrest \citep[for a review and evaluation of such
  hypotheses see][]{Mira1998}.
However, given the existence of female meiotic arrest, 
	and that its timing relative to
	fertilization as well as the interactions between sperm and egg are variable across species  \citep{Karrbook:09} 
	the nature of the meiotic arrest and interactions between sperm and egg may be molded by selection to reduce the opportunity for female meiotic drive, 
	while selfish drivers may evolve in concert, to overcome these adaptations. 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. } 
  }
%However, given that arrest does occur its timing relative to
%fertilization, various features of the interaction between sperm and
%egg and divergence of these features across species  \citep{Karrbook:09} could be subject
%to selection to conflict between female meiotic drivers and alleles
%elsewhere in the genomes of sperm and eggs.
%the mechanism by which it occur
%mechanistic features, and evolutionary changes among taxa
%various features of the
%mechanism of fertilization 
%however, our work does imply that the fertilization requirement does
%not 
%leave female meiosis open to sabotage by sperm collaborators, but
%rather that  an 
%evolutionary loss of female meiotic arrest will result in the loss of sperm defense against female meiotic drivers. }
